---
output:
  pdf_document: default
  html_document: default
---
#Set Seed
In order to obtain the same results

```{r}
set.seed(1964)
```


## HW1.2

*Roll Blue and Red Die,what is the probability that Red Dice < Blue Dice*

#Creation of two independant die -- a red and blue die.
```{r}

red_dice_roll <- function(n){
rolls <- sample(1:6, size = n, replace = TRUE)
return(rolls)}

blue_dice_roll <- function(n){
rolls <- sample(1:6, size = n, replace = TRUE)
return(rolls)}
```




#Determine the probability that the result from the blue die is larger than that of the red die.
```{r}
roll_dice_B_Greater_R <- function(n){
blue_is_greater <- 0
    for (i in 1:n){
if(blue_dice_roll(1) > red_dice_roll(1)){ 
blue_is_greater <- blue_is_greater + 1
}}
  return(blue_is_greater/n)}

## The solution will be equivelent to 
roll_dice_B_Greater_R(50000)
```

#HW 1.3

##One coin is drawn at random (two coins: One is 50% Heads and the second one is 60% Heads)

Coin A = 50%

Coin C = 60%

*If you flip the coin twice and get heads both times what is the probability that this is the fair coin?*


First we must create the functions for the fair and unfair coin 
```{r}
#Function for the fair coin
coin_a <- function(n_flips){
  flips <- sample(c(1,0), size = n_flips, replace = TRUE, 
                  prob = c(0.5, 0.5)) ##Fair chance for heads
        return(flips)
}

#Function for the unfair coin (60% heads bias)
coin_c <- function(n_flips){
  flips <- sample(c(1,0), size = n_flips, replace = TRUE, 
                  prob = c(0.6, 0.4)) ##Unfair chance for Heads
        return(flips)
}
```

Second we must construct another function to count the number of heads obtained in a coin flips alongside a probability for the chance that the coin is the fair coin if Heads is seen twice (HH).

```{r}


#Function to detmermine of achieving Heads Heads
fair_coin_HH <- function(n_trials){
HH_on_flip <- 0
  for(i in 1:n_trials){ 
    if(coin_a(1) + coin_a(1) == 2){
      HH_on_flip <- HH_on_flip +1
    }}
  return(HH_on_flip/n_trials)}


prob_of_fair_coin = fair_coin_HH(100000) * 0.5 
#getting two heads with a fair coin multiplied by the chance that the fair coin is selected
prob_of_fair_coin


```

*After flipping the coin twice and seeing two heads, what is the probability that the next toss is also heads?*

```{r}
#Addition of the probabilities of the next flip being heads
prob_next_flip_is_heads <-
  prob_of_fair_coin*0.5 + (1-prob_of_fair_coin)*0.6

prob_next_flip_is_heads
```







#HW1.4

*Let A and C be events where . . .*

###P(A) = 0.4

###P(A|C)= 0.1

##Explain why it is not possible to have P(C) = 0.8 -- more specifically can you prove an upper or lower bound on P(C)?


P(A|C) = P(A,C) / P(C) = P(C,A) / P(C)

 = > P(C,A) = P(A|C)*P(C)


P(C|A) = P(C,A) / P(A) ==> P(C,A) = P(C|A)*P(A)

.'. 


#HW1.5


