---
output:
  pdf_document: default
  html_document: default
---
#Set Seed
In order to obtain the same results

```{r}
set.seed(1964)
```


## HW1.2

*Roll Blue and Red Die,what is the probability that Red Dice < Blue Dice*

#Creation of two independant die -- a red and blue die.
```{r}

red_dice_roll <- function(n){
rolls <- sample(1:6, size = n, replace = TRUE)
return(rolls)}

blue_dice_roll <- function(n){
rolls <- sample(1:6, size = n, replace = TRUE)
return(rolls)}
```




#Determine the probability that the result from the blue die is larger than that of the red die.
```{r}
roll_dice_B_Greater_R <- function(n){
blue_is_greater <- 0
    for (i in 1:n){
if(blue_dice_roll(1) > red_dice_roll(1)){ 
blue_is_greater <- blue_is_greater + 1
}}
  return(blue_is_greater/n)}

## The solution will be equivelent to 
roll_dice_B_Greater_R(50000)
```

#HW 1.3

##One coin is drawn at random (two coins: One is 50% Heads and the second one is 60% Heads)

Coin A = 50%

Coin C = 60%

*If you flip the coin twice and get heads both times what is the probability that this is the fair coin?*


First we must create the functions for the fair and unfair coin 
```{r}
#Function for the fair coin
coin_a <- function(n_flips){
  flips <- sample(c(1,0), size = n_flips, replace = TRUE, 
                  prob = c(0.5, 0.5)) ##Fair chance for heads
        return(flips)
}

#Function for the unfair coin (60% heads bias)
coin_c <- function(n_flips){
  flips <- sample(c(1,0), size = n_flips, replace = TRUE, 
                  prob = c(0.6, 0.4)) ##Unfair chance for Heads
        return(flips)
}
```

Second we must construct another function to count the number of heads obtained in a coin flips alongside a probability for the chance that the coin is the fair coin if Heads is seen twice (HH).

```{r}


#Function to detmermine of achieving Heads Heads
fair_coin_HH <- function(n_trials){
HH_on_flip <- 0
  for(i in 1:n_trials){ 
    if(coin_a(1) + coin_a(1) == 2){
      HH_on_flip <- HH_on_flip +1
    }}
  return(HH_on_flip/n_trials)}

#Getting two heads with a fair coin multiplied by the chance that the fair coin is selected
prob_of_fair_coin = fair_coin_HH(100000) * 0.5 
prob_of_fair_coin


```

*After flipping the coin twice and seeing two heads, what is the probability that the next toss is also heads?*

```{r}
#Addition of the probabilities of the next flip being heads
prob_next_flip_is_heads <-
  prob_of_fair_coin*0.5 + (1-prob_of_fair_coin)*0.6

prob_next_flip_is_heads
```







#HW1.4

*Let A and C be events where . . .*

###P(A) = 0.4

###P(A|C)= 0.1

##Explain why it is not possible to have P(C) = 0.8 -- more specifically can you prove an upper or lower bound on P(C)?


P(A|C) = P(A,C) / P(C) = P(C,A) / P(C)

 = > P(C,A) = P(A|C)*P(C)


P(C|A) = P(C,A) / P(A) ==> P(C,A) = P(C|A)*P(A)

###Not sure how to proceeed.

#HW1.5

A dianostic test for a disease return one of the following answers; Positive (Has the Disease) and Negative (does not have the disease)

*The probabilities of the test outcomes are as follows*

##For a patient with the disesase. 0.95 Positive / 0.05 Negative
## For a patient that does not have the disease. 0.10 Positive / 0.90 Negative

###*Assume 10% of the popualtion has the disease. If everyone is tested, what proportion of the test results are positive?*


```{r}
#Probabilities of achieving a positive result whether or not you have the disease
with_disease_positive_result = 0.1 * 0.95
without_disease_positive_result = 0.9 * 0.1

prob_of_positive_result = with_disease_positive_result + without_disease_positive_result
prob_of_positive_result
```

##*For a patient who gets a positive result, what is the probability of having the disease*

Definition of conditional probability:

P(A|C) = P(A,C) / P(C) = [P(A)*P(C|A)] / P(C)

P(A): Having the Disease
P(C): Getting a positive result


```{r}
#Known Terms
with_disease_positive_result = 0.1 * 0.95
  without_disease_positive_result = 0.9 * 0.1

  #P(A)    
prob_having_disease = 0.1

#P(A^C)
prob_not_having_disease = 0.9
##########################################################################################################

#P(C)      
prob_getting_positive_result =  
  without_disease_positive_result +
            with_disease_positive_result
#P(A,C)
with_disease_positive_result

#P(A|C) Gets a positive result - prob of having the disease

prob_pos_res_and_having_dis = with_disease_positive_result / prob_getting_positive_result
prob_pos_res_and_having_dis
```


*Now let d be the prevelance of the disease in the population*

*Everyone in the population is screeened for the disease. For each person who got a positive result they are brought back for an additional screening. 25% of the returning population test positive. What is d?*

Q : Second Test is Positive
A : First Test is Positive

#Probability that you get both tests positive

$P(Q|A) = 0.25$

#Probability of getting a positive result times the probability of geeting a positive result

$P(Q,A) = d(0.95)^{2} + (1-d)(0.1)^{2}$

#Probability of getting a positive result (1st test)

$P(A) = 0.95d + (1-d)(0.1)$

$P(Q|A) = P(Q,A) / P(A)$

$P(Q|A) = [d(0.95)^{2}+(1-d)(0.10)^{2}] / [d(0.95) + (1-d)(0.1)]$

$P(Q|A) = [d(0.9025) + (1-d)(0.01)] / [d(0.95) + (1-d)(0.1)]$

$P(Q|A) = [d(0.9025) + (1-d)(0.01)] / [d(0.95) + (1-d)(0.1)]$

$P(Q|A) = [0.9025d + 0.01 - 0.01d] / [0.95d + 0.1 - 0.1d]$

$P(Q|A) = [0.925d + 0.01] / [0.85d + 0.1]$

$d = 0.02105$
