---
title: "Homework 5"
author: "Justin Rigby"
date: "April 28, 2018"
output: html_document
---
  
```{r setup, include=TRUE}
set.seed(1984)
  library(tidyr)
    library(expm)
#Base pair frequencies
pi_A <- 0.1
  pi_C <- 0.2 
    pi_G <- 0.3
      pi_T <- 0.4
```
  
  
1.    
```{r}
P_AGCC <- (pi_A * pi_G * pi_C * pi_C)
  P_AGCC
```
  
  
  
2. 
```{r}

#Function reflecting the Jukes-Cantor equation
JC_ii <- function(b){
p_ii <- (1/4) + (3/4)*(exp(-b*(4/3)))
return(p_ii)  
}

JC_ij <- function(b){
p_ij <- (1/4) - (1/4)*(exp(-b*(4/3)))
return(p_ij)  
}

JC_ij(0.5)
```
  
3.  
  
$ \pi_A = \pi_T = \pi_G = \pi_C = \pi_i = 0.25$  
  
```{r}
pi_i <- 0.25
  n_states <- 4

P_AA <- JC_ii(0.5)*0.25
  P_CA <- JC_ij(0.5)*0.25
    P_GA <- JC_ij(0.5)*0.25
      P_TA <- JC_ij(0.5)*0.25

P_iA <- (P_AA+P_CA+P_GA+P_TA)/(n_states)
P_iA
```
  
4.  
```{r}
```
  
The diagnoles in the P_0.5 matix are the probability that state i stays at state i. The reason why they are so much larger than the other values in the row is because it is more likely for the base pair not to mutate than to mutate. 
  
5.  
That cell represets the probability that the 'A' state remains in the 'A' state. As mentioned above, i is more likely to stay as i because the chances of a mutation at specific basepair is low. This is further explained by the original R matrix which shows that the 1st row has the lowest values for the rates.    
  
6.  
  
```{r}
p_GA <- 0.028 
  p_GT <- 0.251

#Probability that it goes from G to A
P_GA <- p_GA*pi_G
P_GA
    
#Probability that is goes from G to T
    
P_GT <- p_GT*pi_G
P_GT    
```
It is true becuase the frequency at which specific base pairs are mutated varies depending on the base pair (e.g $\pi_A = 0.1 ~~\pi_C = 0.2 ~~\pi_G = 0.3 ~~\pi_T = 0.4$). These parameters are integrated into the matrix via a scaling calculated from the base pair frequencies.   

7.  
  
```{r}
p_AC <- 0.037 
  p_AT <- 0.148

#Probability that it goes from G to A
P_AC <- p_AC*pi_A
P_AC
    
#Probability that is goes from G to T
    
P_AT <- p_AT*pi_A
P_AT    
```
It is true becuase the frequency at which specific base pairs are mutated varies depending on the base pair (e.g $\pi_A = 0.1 ~~\pi_C = 0.2 ~~\pi_G = 0.3 ~~\pi_T = 0.4$). These parameters are integrated into the matrix via a scaling calculated from the base pair frequencies.   
  
8.  

```{r}
library(expm)

matrix_Q <- matrix(c(-0.641,0.0321, 0.0641, 0.0962, 
                     0.064, -1.058, 0.256, 0.321, 
                     0.192, 0.385, -1.09, 0.577, 
                     0.385, 0.641, 0.769, -0.944
                     ), nrow = 4, ncol = 4)
matrix_b_0.5 <- expm(matrix_Q*0.5)
matrix_b_50 <- expm(matrix_Q*50)
  matrix_b_50
#Ratios of the results
matrix_b_50/25
```
  
The numbers have begun to converged to the ratio of the starting frequencies 
(e.g $\pi_A = 0.1 ~~\pi_C = 0.2 ~~\pi_G = 0.3 ~~\pi_T = 0.4$). 
  
9.  
$P_{AC} = \pi_{A} * p_{AC} = 0.037 * 0.1 = 0.0037~~~~|~~~~P_{CA} = \pi_{C} * p_{CA} = 0.2 * 0.018 = 0.0036~~~~|$
  
If this were to be a model where the parameters were equal for 
all outcomes (e.g. $\pi_{A} = \pi_T = \pi_G =\pi_C =  0.25$) we would see a complete reversal of the number -- meaning 
they would be equal. The reason why there was a non-reversible outcome was becuase there was not equal frequencies
for all basepairs.  
  
  
  
10.  
##  
a.  
```{r}
matrix_Q <- matrix(c(-0.641,0.0321, 0.0641, 0.0962, 
                     0.064, -1.058, 0.256, 0.321, 
                     0.192, 0.385, -1.09, 0.577, 
                     0.385, 0.641, 0.769, -0.944
                     ), nrow = 4, ncol = 4)

matrix_b_0.8 <- expm(matrix_Q*0.8)
  matrix_b_0.8
```

b.  
$P_{TG|50} = \pi_{T} * p_{TG|50} = 0.4* 0.2450571 = 0.09802284$  
$P_{TG|0.5} = 0.188 * 0.4 = 0.0752$  
The transition from T to G is more probable in the 0.8 than the 0.5 branch because it is longer and has a higher
probability of having substitutions.  
   
c.  
  
```{r}
P_AA <- pi_A * 0.61003692
  P_AC <- pi_A * 0.05976266
    P_AC <- pi_A * 0.05976266
      P_AT <- pi_A * 0.2092071
      
P_ACCT <- P_AA*P_AC*P_AC*P_AT
P_ACCT
```
  
d.  
  
```{r}
P_CA <- pi_C * 0.02992124
  P_CC <- pi_C * 0.48153010
    P_CC <- pi_C * 0.48153010
      P_CT <- pi_C * 0.2973463
      
P_ACCT <- P_CA*P_CC*P_CC*P_CT
P_ACCT

```
  
e.  
```{r}
n_states <- 4
P_AA <- pi_A * 0.61003692
  P_CA <- pi_C * 0.02992124
    P_GA <- pi_G *  0.04193268
      P_TA <- pi_T * 0.05230103
P_iA <- (P_AA+P_CA+P_GA+P_TA)/n_states
P_iA      
```
  
f.  
```{r}
n_states <- 4
P_AC <- pi_A * 0.05976266
  P_CC <- pi_C * 0.48153010
    P_GC <- pi_G *  0.13200636
      P_TC <- pi_T * 0.14881736
P_iC <- (P_AC+P_CC+P_GC+P_TC)/n_states
P_iC
```
  
g.  
  
```{r}
#Probability of the X sequence
P_iX <- ((pi_A * matrix_b_0.5[1,1] + 
            pi_C * matrix_b_0.5[2,1] + 
            pi_G * matrix_b_0.5[3,1] + 
            pi_T * matrix_b_0.5[4,1])/4) * 
          ((pi_A * matrix_b_0.5[1,2] + 
            pi_C * matrix_b_0.5[2,2] + 
            pi_G * matrix_b_0.5[3,2] + 
            pi_T * matrix_b_0.5[4,2])/4) *
          ((pi_A * matrix_b_0.5[1,4] + 
            pi_C * matrix_b_0.5[2,4] + 
            pi_G * matrix_b_0.5[3,4] + 
            pi_T * matrix_b_0.5[4,4])/4)
P_iX

#Probability of the Y sequence
P_iY <- ((pi_A * matrix_b_0.8[1,3] + 
            pi_C * matrix_b_0.8[2,3] + 
            pi_G * matrix_b_0.8[3,3] + 
            pi_T * matrix_b_0.8[4,3])/4) * 
          ((pi_A * matrix_b_0.8[1,1] + 
            pi_C * matrix_b_0.8[2,1] + 
            pi_G * matrix_b_0.8[3,1] + 
            pi_T * matrix_b_0.8[4,1])/4) *
          ((pi_A * matrix_b_0.8[1,4] + 
            pi_C * matrix_b_0.8[2,4] + 
            pi_G * matrix_b_0.8[3,4] + 
            pi_T * matrix_b_0.8[4,4])/4)
P_iY

#The chance of getting both sequences derived from the same anscestor.
P_XY <- P_iY*P_iX

P_XY

```
  
11.  
  
a.  

$\pi_G = \pi_C = \frac{3}{8} = 0.375$  
$\pi_A = \pi_T = \frac{1}}{8} = 0.125$  
  
b.  
  
```{r}
R_JT14 <- matrix(c(0, 1, 2, 3,
                 1, 0, 4, 5,
                 2, 4, 0, 6,
                 3, 5, 6, 0), ncol = 4, nrow = 4)
R_JT14
```
  
c.  
  
```{r}
Q_unscaled <- function(pi_A, pi_C, pi_G, pi_T, RT_JT14){

R_JT14 <- matrix(c((R_JT14[,1]*pi_A), 
                 (R_JT14[,2]*pi_C),
                 (R_JT14[,3]*pi_G),
                 (R_JT14[,4]*pi_T)), ncol = 4, nrow = 4)
R_JT14[1,1] <- (-1*(sum(R_JT14[1,]))) 
  R_JT14[2,2] <- (-1*(sum(R_JT14[2,]))) 
    R_JT14[3,3] <- (-1*(sum(R_JT14[3,])))
      R_JT14[4,4] <- (-1*(sum(R_JT14[4,]))) 

return(R_JT14)}

Q_Un_JT14 <- Q_unscaled(pi_A, pi_C, pi_G, pi_T, RT_JT14)
  Q_Un_JT14


```
  
d.  
  
```{r}
pi_A <- 0.125
  pi_C <- 0.375
    pi_G <- 0.375
      pi_T <- 0.125
      
scale_fx <- function(matrix, pi_A, pi_C, pi_G, pi_T){
scale <- (-1/(matrix[1,1]*pi_A + 
                matrix[2,2]*pi_C + 
                  matrix[3,3]*pi_G + 
                    matrix[4,4]*pi_T))
return(scale)
}

scale <- scale_fx(Q_Un_JT14, pi_A, pi_C, pi_G, pi_T)
```
  
e.  
```{r}

Q_JT14 <- (Q_Un_JT14*scale)
  Q_JT14
```
  
f.  
  
```{r}
P_0.5 <- expm(Q_JT14*0.5)
  P_0.5

P_20 <- expm(Q_JT14*20)
  P_20
```
  