---
title: "Networks_Lab_4_Rigby"
author: "Justin Rigby"
date: "April 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  library(bipartite)
    library(ggplot2
            )
```
  
  
  
1.  Try comparing to a fixed shuffling algorithm like curveball. Are there differences in the results when using different null models? Why?
  
```{r}
# Loading data into R from sample dataset
data("memmott1999")
#Creation of an un-weighted binary matrice of the data 
memmott_bin <- 1*(memmott1999>0)
nodf <- nestednodf(memmott_bin)
  nodf
    plot(nodf)

#Using a fixed shuffling algorithm (curveball) to see if there are different results 
nodf_eval_2 <- oecosimu(memmott1999_binary, nestednodf, "curveball", nsimul = 100)
nodf_eval_2$statistic # This is the NODF of the observed network
names(nodf_eval_2$oecosimu) # look at the ?oecosimu help for details on what are these values.
nodf_eval_2 # The networks is significantly nested.
```
From the results we can see that there is a change in the significant difference of the N.Rows and NODF when compared to the null (curveball). When the r00 was ran we could see that there was significant differences between the 'r00' null and the empirical data, but when 'curveball' null is used it does not show any significance in the N.rows and NODF groups.  
  
The differences between the two null models may actually explain why there is a difference in significance of the data. The 'r00' null model is an algorithm which creates random matrices using only the size of the matrix, and the number of filled locations as parameters to building new matrices. Meaning, theses matrices are randomized and not directly related to the empirical data. The 'curveball' algorithm instead chooses two separate rows in the matrix and finds the set of unique species interactions in one of the rows but not the other. These selected taxa are then distributed across different rows preserving the original row frequencies. The 'curveball' algorithm is also found to pertube the matrices more than other methods allowing for a conservative model, and it is alsoa model based on the input sequences and not a randomized matrix based on just the number of interactions and the size of the beginning matrix.    
  
2. Nestedness has been shown to be correlated with network density and the number of species. That rencers comparisons of nestedness scores across many network statistically inappropriate. How can we comapre valuse of nestedness across networks?  
  
We use z-score transformations of the data in order to allow us to compare otherwise incomparable data sets.  
  
  
3. What can you say about the tolerance of the network to difference extinction regimes? REpeat that excersise, now removing pollinators. Is the tolerance of hte network diffrerent?  
  
2.1
  
Pseudocode  
#----------------------------------------------------------------------------------------------
1.Choose node  
2.identify non-zero entries  
LOOP{  
3. Set entry to zero  
4. Calculate sum of row 
  If zero, plant is extinct  
}  


```{r}
matrix <-  memmott1999_binary
max(names(colSums(matrix)))
colSums(matrix)

```
  


  
  
  