---
output:
  pdf_document: default
  html_document: default
---

#HW2.1

```{r}
set.seed(1984)
```

In the posted homework solutions, there is a 20% chance that there is an error in the posted solutions. If an error is present, then the number of students who send emails about the error follows a poisson(3) distrubution.
  
*If no emails were recieved, what is the probability that there was no error*

####We must first define our information

The probability of an error in the posted solutions is 20%  
$\therefore$ $prob.of.error.hw = P(E) = 0.20$  
  
Also,  
$\therefore$ $prob.of.no.error.hw = P(E^{c}) = 0.80$ 


```{r}
set.seed(1984)
hw_errors <- function(n){
  num_of_emails.df <- data.frame() #Creation of Empty Data Frame
for(i in 1:n){x_ind <- 0 #Indicator Variable
  x_ind <- sample(c(0,1),1,replace = TRUE, prob = c(0.8,0.2)) #Probability of Error in HW
if(x_ind == 1){num_of_emails.df <- rbind(num_of_emails.df, rpois(1,3))}
} #If there is an error what is the number of emails received with  
#respect to the poisson distrubution
no_emails_with_error <- sum(num_of_emails.df == 0) #Sum the number of 0 email occurances
return(1-(no_emails_with_error/n)) #Sum of the number of 0 emails weighted by the total  
##number of HW Solutions posted
    }

#Therefore, the probability of an error being in the solutions, yet no email   
#received about it, is about  
hw_errors(50000)
```


#HW2.2

Calculate P(X is odd) in each setting below.  
  (a) X ~ Geometric(0.7)  
  (b) X ~ Binomial(101, 0.5)  
  (c) First let Y ~ N(0,1), and then let X be the answer you get when you round Y to the nearest integer.  
###########################################################################################

a) X ~ Geometric(0.7)

 
```{r}
#Function to determine if number is odd
is.odd <- function(x){ x%%2!=0
}
n=50000
p=0.7
prob_geom_odd <- function(n, p){
geom.df <- data.frame(rgeom(n, p))
colnames(geom.df) <- c("Binary")
geom.df <- (geom.df%%2!=0)
geom.binary.df < (1*geom.df)

return(sum(geom.binary.df)/n)
}

prob_geom_odd(500000, 0.7)
```


b) Binomial(101, 0.5)

```{r}
prob.binom.odd <- function(n, size, prob){
#Creation of n size draws from a normal distrubution divided by 2 to indicate odd variables
  binom.df <- data.frame(rbinom(n, size, prob))
    colnames(binom.df) <- c("Binary")
    #Since only even numbers will be integers, this is to test if they are integers  
      binom.binary.df <- (1 *(binom.df%%1==0))

return(1 - (sum(binom.binary.df)/n))  
}

prob.binom.odd(101, 500000, 0.5)
```


(c) First let Y ~ N(0,1), and then let X be the answer you get when you round Y to the nearest integer. 
```{r}
prob_norm_odd <- function(n){
  norm.df <- data.frame(rnorm(n))
    norm.round.df <- round(norm.df)
      colnames(norm.round.df) <- c("Binary")
norm.round.binary.df <- (1 *(norm.round.df%%2!=0))
return(1 - (sum(norm.round.binary.df)/n))  
  }
prob_norm_odd(50000)
```
