#HW2.1

```{r}
set.seed(1984)
```

In the posted homework solutions, there is a 20% chance that there is an error in the posted solutions. If an error is present, then the number of students who send emails about the error follows a poisson(3) distrubution.
  
*If no emails were recieved, what is the probability that there was no error*

####We must first define our information

The probability of an error in the posted solutions is 20%  
$\therefore$ $prob.of.error.hw = P(E) = 0.20$  
  
Also,  
$\therefore$ $prob.of.no.error.hw = P(E^{c}) = 0.80$ 


```{r}
set.seed(1984)
hw_errors <- function(n){
  num_of_emails.df <- data.frame() #Creation of Empty Data Frame
for(i in 1:n){x_ind <- 0 #Indicator Variable
  x_ind <- sample(c(0,1),1,replace = TRUE, prob = c(0.8,0.2)) #Probability of Error in HW
if(x_ind == 1){num_of_emails.df <- rbind(num_of_emails.df, rpois(1,3))}
} #If there is an error what is the number of emails received w/ respect to the poisson distrubution
no_emails_with_error <- sum(num_of_emails.df == 0) #Sum the number of 0 email occurances
return(no_emails_with_error/n) #Sum of the number of 0 emails weighted by the total number of HW Solutions posted
    }

#Therefore, the probability of an error being in the solutions, yet no email received about it, is about. . .  
hw_errors(50000)
```


#HW2.2

Calculate P(X is odd) in each setting below.  
  (a) X ~ Geometric(0.7)  
  (b) X ~ Binomial(101, 0.5)  
  (c) First let Y ~ N(0,1), and then let X be the answer you get when you round Y to the nearest integer.  
###########################################################################################

a) X ~ Geometric(0.7)

 
 
 